---
title: "Logistic Regression"
author: "Gabriel Costa"
date: '2022-07-15'
output:
  md_document:
    variant: markdown_github

---

---

Using data from Yelp to predict the probability of a restaurant having an above average rating.

---

##### 1. Importing required libraries and data set.

Reference to the original data: https://www.yelp.com/dataset

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(caret)
```


```{r}
yelp <- read.csv("https://raw.githubusercontent.com/ggc0sta/tmu/main/data/yelp_dataset.csv",
                 stringsAsFactors = TRUE)
yelp <- yelp[,!(names(yelp)  %in% c("id"))] #removing id column
```

---

##### 2. Describing data

* **city** - Location of each restaurant. Toronto, Markham or Mississauga
* **class** - Above average rating (0 = no, 1 = yes) 
* **review_count** - number of reviews in yelp
* **categories** - type of restaurant

```{r}
summary(yelp) #no missing values
```


```{r}
head(yelp)
```

---

##### 3. Preparing the data for modeling.
```{r}
set.seed(1) #for reproducibility
sampleSize <- floor(0.8*nrow(yelp))
trainIndex <- sample(1:nrow(yelp), size = sampleSize)

train <- yelp[trainIndex,]
test <- yelp[-trainIndex,]

# set sizes
dim(train);dim(test)
```

##### 4. Creating a Logistic Regression model.

The **glm** function converts the categorical variables **city** and **category** into dummies. As the factors are ordered alphabetically, *Markham* is used as the base city and *Asian* as the base restaurant type.

```{r}
levels(yelp$city)
levels(yelp$categories)
```

```{r}
model1 <- glm(formula = class ~ ., data =  train, 
              family = "binomial")

summary(model1)
```
---

##### 5. Predicting the probability of a **coffee shop** in **Toronto** to be ranked as above average (**class** = 1), expecting to receive 200 reviews.
```{r}
# predict function
predict(model1, 
        newdata=data.frame(city = "Toronto", 
                           review_count = 200, 
                           categories = "Coffee or Sandwiches"), 
        type="response")

# manual calculation
b0 <- model1$coefficients["(Intercept)"][[1]]
b1 <- model1$coefficients["cityToronto"][[1]]
b2 <- model1$coefficients["review_count"][[1]]
b3 <- model1$coefficients["categoriesCoffee or Sandwiches"][[1]]

a <- b0 + b1 + b2*200 + b3
exp(a)/(1+exp(a))

```

---

##### 6. Evaluating the performance of the model based on accuracy.

```{r}
predictions <- predict(model1, 
                       newdata = test, 
                       type="response") 
predictions <- ifelse(predictions > 0.5, 1, 0)
confusionMatrix(data = as.factor(predictions), reference =  as.factor(test$class))

```



*Based on CMTH642 - Advanced Data Analytics, Lab 7*